---
title: "Creating graphs with ggplot"
author: "Mies Casey"
date: "1/27/2020"
output: html_document
---
# Plots with ggplot2

We are going to make plots in R using the ggplot2 packages from the tidyverse.

-tidyverse: suite of packages for wrangling & viz ( ggplot2)
-readxl: to read in excel files
- here: to help with filepaths
```{r}
install.packages("tidyverse")
```



```{r setup, message=FALSE} 
library(tidyverse) # install.packages ("tidyverse")
library(readxl) # install.packages ("readxl")
library(here) # install.packages ("here")
```

# California National Parks Data

```{r, message = FALSE}

ca_np <- read_csv(here("data/ca_np.csv"))
ci_np <- read_xlsx(here("data/ci_np.xlsx"))
```

### Make my own first graph

We're going to make a lineplot of year (x-variable) vs vistitors (y-variable) in ggplot:
command + option + i 

```{r}
ggplot(data = ci_np, aes( x = year, y = visitors)) + 
  geom_col()
```

# Update the geom_ to geom_point(), geom_col(), geom_area()

Making a base layer for my channel island plots:

```{r}
gg_base <- ggplot(data = ci_np, aes( x = year, y = visitors))

gg_base + geom_col()
```

### Now Let's Start Customizing!

- Line / point colors: color
- Change the line type: linetype
- Change point style: pch
- change transparency: alpha

```{r}
gg_base + 
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

```{r}
gg_base +
  geom_point(
    color = "dodgerblue",
    pch = 19,
    size = 2
  )
```

### Update graph aesthetics based on variables

```{r}
gg_base +
  geom_point(
    aes(size = visitors, 
        color = year),
    pch = 16
  )
```
# Add an argument to update the color based on the year value
