---
title: "Lobster Data"
author: "Miles Casey"
date: "1/27/2020"
output: html_document
---

```{r}
library(tidyverse) #install.packages("tidyverse")
library(readxl)
library(here)
#install.packages("skimr")
library(skimr) # install.packages("skimr")
#install.packages("kableExtra")
library(kableExtra) # install.packages("kableExtra)

# read in lobster data
lobsters <- Lobsters <- read_xlsx(here("data", "lobsters.xlsx"),
                                        skip = 4)
```

Viewing our data with skimr

```{r}
skimr::skim(lobsters) #explicitly stated from skimr
skim(lobsters)
```

# Pivot Tables
In R, Pivot tables are : `group_by() %>% summarize()`
cmd + shift + m
letter "n" will count observations
? + function brings up help menu
```{r}
# group by year
lobsters %>%
  group_by(year) %>% 
  summarize(count_by_year = n())

# group by year and site
lobsters %>%
  group_by(year,site) %>% 
  summarize(count_by_year_site = n())

# group by site and year
lobsters %>%
  group_by(site, year) %>% 
  summarize(count_by_site_year = n())
```

Summarize by multiple variables
na.rm removes all data sets that were missed or left out to make sure there are no blank values
assigning to an object allows for easily re-using code you already have written
```{r}
siteyear_summary <- lobsters %>%
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE), 
            sd_size_mm = sd(size_mm, na.rm = TRUE))
```

Table formatting with `kableExtra::kable()`
```{r}
siteyear_summary %>% 
  kable()
```

There are `r nrow(lobsters)` total lobsters included in this report.

## Activity
1. Build from our analyses: calculate the median lobster size for each site and year (hint: `size_mm` and a function called `median()`
2. Sync this to github

```{r}
siteyear_summary <- lobsters %>%
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            median_size_mm = median(size_mm, na.rm = TRUE), 
            sd_size_mm = sd(size_mm, na.rm = TRUE))
ggplot(data = siteyear_summary, aes(x = year, y = median_size_mm, color = site)) +
  geom_line()
ggsave(here("figures", "lobsters-plot.png"))
```


